Rotation effects on the Lyman-$\alpha$ line morphology in distant galaxies

### 2013-06-11 paper/Analisis.tex
text in the escape fraction subsection

\usepackage[normalem]{ulem}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

\subsection{Line width and peak maxima}
\label{sec:widthpeak}
\label{fig:differentvelocities}

@@

@@

The analytic expectation for the slab with homogeneously emitted
sources is $\langle N_{\rm  scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}, a factor of $0.72$ lower than the case of the
centrally emitted photons. In our case we find that for the static
spheres this factor is $XXX$.

%%%%%

In order to gain a deeper understading of these results we 2D
histograms for the number of scaterrings as a function of the outgoing
dimensionless frequency $x$. In Figure \ref{fig:Nscatt2D} we show
the results of such histogram in the case $\tau{\rm H}=10^5$ for the
static case and $V_{\rm max}=300$\kms. The upper (lower) panels show the
results for the homogeneous (central) source distribution. The color
scale is logarithmic in the number of photons at a certain value
$x-N_{\rm scatt}$. This Figure clearly support our hypothesis about
the photosphere in the homogeneous distribution, in this case most of
the photons that left with $x\sim 0$ have escaped with less than $10$
scatterings, clearly explaining the origin of a single central
peak. However, for a central distribution the situation is
different. In this case the number of scatterings remains high, on the
order of the optical depth, but the two peaks do get closer to each
other. In this situation each scattering leaves the photon closer to
the center of the line than the static case.

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Log of the Average number of scatterings of the outgoing
  photons as function of the velocity. \label{fig:Nscatt}}  
\end{figure}

\begin{figure*}
     \includegraphics[width=0.45\textwidth]{2dHistogram0t5HOM.png}
     \includegraphics[width=0.45\textwidth]{2dHistogram300t5HOM.png}    \\
     \includegraphics[width=0.45\textwidth]{2dHistogram0t5.png}
     \includegraphics[width=0.45\textwidth]{2dHistogram300t5.png}    
    \caption{2D histogram of $N_{scatt}$ vs $x$. The upper (lower) pannels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and the right
      $V_{max}=300km/s$. The colour scale is locarithmic\label{fig:Nscatt2D}}  
\end{figure*}

%%%%
\subsection{Escape Fraction}
\label{sec:EF}

Until now we have studied models in absence of dust, but is known (REFERENCE) 
that there is presence of dust in high redshift galaxies. Our models with 
dust are treated as is explained in detail in  the appendix of (Forero et al 2011).

Of particular interest is to compute the escape fraction of Lyman $\alpha$ 
photons coming from the most distant galaxies, due to the fact that with the 
observed intensity of the Lyman alpha line quantities as the LF and SFR can 
be derived (put some references).

Previous studies have shown the correlation of the Escape fraction with galactic 
mass (Laursen et al 2009, Dayal et al 2010) abundances and the kinematics of dust. 
In order to study pure rotational effects in the escape fraction we fixed the dust 
abundance $\tau_{A}=0.01$ and the galaxy mass. We compute the escape fraction for 
the models described in Table \ref{table:models}.

For a realistic model we also take into account the viewing angle, first we fixed 
the viewing angle at $\theta = 0$  Fig.~\ref{figure:efvsv} and then we fixed the 
velocity in order to see the escape fraction correlation with the viewing angle. 
Therefore we define the escape fraction as:\\

\begin{equation}
F_{e}=\dfrac{\Sigma_{NI} \vec{k}\cdot \vec{o}}{\Sigma_{NF}\vec{k}\cdot \vec{o}}
\end{equation}

Where NI is the initial number of photons and NF is the final,
$\vec{k}$ is the rotation axis direction and ${o}$ the observer
direction. With this definition we compute the escape fraction for all
of our models, the results are presented in Fig \ref{figure:efvsv}\\

\begin{figure}
  \includegraphics[width=0.45\textwidth]{EscapeFraction.png}
   \caption{Escape fraction for all the models. continious lines represent the homogenous models, while dashed lines represent the central models}\label{figure:efvsv}
\end{figure}

In Figure \ref{figure:efvsv} we found that in the central distribution
the escape fraction does not change with velocity while it does in the
optical depth (See Verhamme 2006, an argument about this). On the other
hand for the homogeneous distribution we
found that for higher velocities photons escape easily. The difference
between this two results rely in the fact that in the homogeneous
distribution photons are emitted closer to the escape surface and this
makes this configuration more sensitive to rotation while in the
central configuration the escape fraction depends mainly in the amount
of gas rather than in rotation.

As a final test we compare our results with the analytical solution of 
the slab developed by (Neufeld 1990) Fig. ~\ref{figure:efvsNeufeld}, as 
the geometry we use is different from the one described for Neufeld, we 
dont expect the same results, in fact we found that for the homogenous 
sphere the escape fraction is higer than the slab (see ref, mark). Also 
we notest that for $\tau<10^{6}$ the escape fraction does not increase 
as it will be expected, This is due to the fact that the condition 
$a\tau_{h}>>\tau_{d}$ in not valid any more.

\begin{figure}
  \includegraphics[width=0.45\textwidth]{Neufeld.png}
 \label{figure:efvsNeufeld}\caption{Escape fraction of this work
   compared with Neufeld analytical solution}  
\end{figure}

\subsection{The Case of Asymmetric Emission}

As we know there is unlikely to find galaxies with radiation sources 
distributed homogeneously. Most of them are in a clumpy distribution 
(Laursen et al 2013*) which affected the resulting spectra. In order 
to study an inhomogeneous distribution we set up a model in which we 
select certain photons that are placed in a specific place but that 
are not symmetrically distributed Fig.~\ref{figure:distributions} shows
the distributions we set up.

\begin{figure}
  \includegraphics[width=0.40\textwidth]{Distribution.png}
  \caption{Inhomogeneous distributions of 
 photons, Blue area represents the photons in distribution 1 while red
 area are the selected photons for distribution 2. The arrow points to
 the observer position\label{figure:distributions}}  
\end{figure}

In Fig.~\ref{figure:inhomogeneous} we show the resulting spectra of 
distribution 1 and 2, the first effect we see is the asymmetry of the
double peak, in the homogeneous and central distribution we see double
peaks with the same height, while in this case one peak is higher than 
the other. For $\tau=10^{5}$ we found that in distribution 1 the blueshifted
peak is higher than the redshifted, and for distribution 2 the redshifted
peak is the highest.

An other important fact here is the asymmetry of the spectra with respect 
to the line center, in particular photons selected in distribution 1 
present a blueshift in their spectra while photons selected in distribution 
2 present a redshift. This effect becomes stronger as velocity increase 
~\ref{figure:inhomogeneous}

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{InomogeneousModelt5.png}
  \includegraphics[width=0.45\textwidth]{InomogeneousModelt6.png}
 \caption{Inhomogeneous model for velocities $50km/s$ and $300km/s$ ,
   (left) with $\tau=10^5$, (right) with
   $\tau=10^6$\label{figure:inhomogeneous}}  
\end{figure*}

\subsection{Integrated flux in a narrowband filter}

Until now we have shown the main effects of gas and dust rotation in
the Ly$\alpha$ line morphology such as the escape fraction, the FWHM,
the maxima position, also we see that shape of the outgoing spectra
depends on the position of the observer. It is important to see if
rotational effects are detectable in observational methods involving
the Ly$\alpha$ line.

One of the most used methods to detect high redshift galaxies using
the Ly$\alpha$ line is using a narrowband selection, we make this
analysis based on the results obtained by Steidel (2011) in this work
(EXPLAIN a lit of bit more about their work) they used tree narrowband
filters for tree different redshifts resumed in
Table~\ref{table:NBfilters}, We want to know how much the integrated
flux change due to rotational effects in this NB filters, for the
models we simulated with CLARA.

\begin{table*}
\begin{center}
\begin{tabular}{cccc}\hline
Model & SSA22a 4980/80   & HS1549 4667/80 & HS1700 4018/90\\
\hline
\end{tabular}
\caption{
Fluxes for tree different narrow band filters.
} 
\label{table:NBfilters}
\end{center}
\end{table*}

In table \ref{table:NBfilters} we present the results of the flux in
every  narrowband filter for the homogeneous model at different
velocities and hydrogen  optical depth $\tau_{0}$. As is well known an
increase in the optical depth makes  the line peaks separation
bigger. In fact for some cases the line width is larger  than the NB
filer width fig.(put ref fig) for those cases we modify the redshift
in the available range in order to make the peak maxima match the NB
filter center.  For all cases we found that as velocity increase the
flux is less.

In the case of the dusty model we found the same trend of the flux
with the velocity  but in this cases the effect is no that strong.  We
also found some dependency with the viewing angle.

\begin{figure}
  \includegraphics[width=0.45\textwidth]{NB7tDifVHOM.png}
 \label{figure:efvsNeufeld}\caption{NB filter located at the center of
   the maxima, with different redshift}  
\end{figure}

@@

... Comparison with Verhamme et al. results on the rotation
... Compare with Kulas et al (Figure 3), Rotation on the lyman alpha
line convert double peak profiles into a single one. comments about
rotation with inflows and outflows.

... The results derived in this paper have consequences on the
interpretation of galaxy observations in the Lyman alpha line.
..compare steidel et al (2011)
% taking into account equivalent widths results with narrow band
% filters we can make a criterium in order to select wich galaxie
% could be seen in actual observations.

>>>>

\usepackage[normalem]{ulem}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

\subsection{Line width and peak maxima}
\label{sec:widthpeak}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Logarithm of the average number of scatterings as function of
  the velocity. \label{fig:Nscatt}}   
\end{figure}

\begin{figure*}
     \includegraphics[width=0.45\textwidth]{2dHistogram0t5HOM.png}
     \includegraphics[width=0.45\textwidth]{2dHistogram300t5HOM.png}    \\
     \includegraphics[width=0.45\textwidth]{2dHistogram0t5.png}
     \includegraphics[width=0.45\textwidth]{2dHistogram300t5.png}

    \caption{2D histogram of $N_{scatt}$ vs $x$. The upper (lower) pannels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and the right
      $V_{max}=300km/s$. The colour scale is locarithmic\label{fig:Nscatt2D}}  
\end{figure*}

@@

The analytic expectation for the slab with homogeneously emitted
sources is $\langle N_{\rm  scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}, a factor of $0.72$ lower than the case of the
centrally emitted photons. In our case we find that for the static
spheres this factor is $xxxx$.

In order to gain a deeper understanding of these results we 2D
histograms for the number of scaterrings as a function of the outgoing
dimensionless frequency $x$. In Figure \ref{fig:Nscatt2D} we show
the results of such histogram in the case $\tau{\rm H}=10^5$ for the
static case and $V_{\rm max}=300$\kms. The upper (lower) panels show the
results for the homogeneous (central) source distribution. The color
scale is logarithmic in the number of photons at a certain value
$x-N_{\rm scatt}$. This Figure clearly support our hypothesis about
the photosphere in the homogeneous distribution, in this case most of
the photons that left with $x\sim 0$ have escaped with less than $10$
scatterings, clearly explaining the origin of a single central
peak. However, for a central distribution the situation is
different. In this case the number of scatterings remains high, on the
order of the optical depth, but the two peaks do get closer to each
other. In this situation each scattering leaves the photon closer to
the center of the line than the static case.

\subsection{Dusty Clouds: Escape Fraction}
\label{sec:escapefraction}

\begin{figure}
  \includegraphics[width=0.45\textwidth]{EscapeFraction.png}
   \caption{Escape fraction as a function of rotational velcity. The
     continous (dashed) lines correspond to homogeneous (central)
     models.
     \label{fig:efvsv}}
\end{figure}

\begin{figure}
  \includegraphics[width=0.45\textwidth]{Neufeld.png}
 \label{fig:efvsNeufeld}\caption{Escape fraction as a function of the
   product $(a\tau_{\rm H})^{1/3}\tau_{a}$. The analytic solution for
   the infinite is slab shown as a continous line. Different dashed
   lines correspond to different rotational velocities.}   
\end{figure}

If there is a change in the number of scatterings with rotational
velocity, there must be a change in the probability of being absorbed
if dust is present. We study how rotation impacts the escape fraction
of \ly photons.

Taking into account the results in the previous section, we do not expect any
change in the case of central distributions given that the average
number of scatterings remain constant. On the other hand, in the
homogeneous distribution there is a considerable drop in the number of
scatterings as $V_{\rm max}$ increases, this preconditions the escape
fraction to increase as well.

Figure \ref{fig:efvsv} shows the dependence of the escape fraction as
a function of rotational velocity. This Figure illustrates how our
reasoning in the previous paragraph is correct. The escape fraction
barely changes for the central source distribution, while it rises
as the velocity $V_{\rm max}$ increases for the homogeneous case. The
velocity has a higher relative impact in the models with low optical
depth $\tau_{\rm H}=10^5$,$10^{6}$, where it can raise from $xxx$ in
the static case up to $xxx$ at $V_{\rm rot}=300$\kms.

In Figure \ref{fig:efvsNeufeld} we put these results in the context of
the analytical solution for the infinite slab \citep{Neufeld90} shown
by the continous line. In this setup the analytic solution depends
solely on the product $(a\tau_{\rm   H})^{1/3}\tau_{a}$. This valid
only in the limits $a\tau_{\rm   H}\gg 1$. The dashed lines in Figure
\ref{fig:efvsNeufeld} corespond to the cases of different
velocities. We observe that in all cases the escape fraction is higher
than in the slab configuration. We also note that for $\tau_{\rm H}=10^{5}$
the escape fraction does not increase as expected, This is
due to the fact that the condition $a\tau_{h}\gg 1$ does not hold.

\subsection{Different line of sights}

\subsection{Off-Centered emission}

%As we know there is unlikely to find galaxies with radiation sources 
%distributed homogeneously. Most of them are in a clumpy distribution 
%(Laursen et al 2013*) which affected the resulting spectra. In order 
%to study an inhomogeneous distribution we set up a model in which we 
%select certain photons that are placed in a specific place but that 
%are not symmetrically distributed Fig.~\ref{figure:distributions} shows
%the distributions we set up.

%\begin{figure}
%  \includegraphics[width=0.40\textwidth]{Distribution.png}
%  \caption{Inhomogeneous distributions of 
% photons, Blue area represents the photons in distribution 1 while red
% area are the selected photons for distribution 2. The arrow points to
% the observer position\label{figure:distributions}}  
%\end{figure}

%In Fig.~\ref{figure:inhomogeneous} we show the resulting spectra of 
%distribution 1 and 2, the first effect we see is the asymmetry of the
%double peak, in the homogeneous and central distribution we see double
%peaks with the same height, while in this case one peak is higher than 
%the other. For $\tau=10^{5}$ we found that in distribution 1 the blueshifted
%peak is higher than the redshifted, and for distribution 2 the redshifted
%peak is the highest.

%An other important fact here is the asymmetry of the spectra with respect 
%to the line center, in particular photons selected in distribution 1 
%present a blueshift in their spectra while photons selected in distribution 
%2 present a redshift. This effect becomes stronger as velocity increase 
%~\ref{figure:inhomogeneous}

%\begin{figure*}
%  \includegraphics[width=0.45\textwidth]{InomogeneousModelt5.png}
%  \includegraphics[width=0.45\textwidth]{InomogeneousModelt6.png}
% \caption{Inhomogeneous model for velocities $50km/s$ and $300km/s$ ,
%   (left) with $\tau=10^5$, (right) with
%   $\tau=10^6$\label{figure:inhomogeneous}}  
%\end{figure*}

%\subsection{Integrated flux in a narrowband filter}

%Until now we have shown the main effects of gas and dust rotation in
%the Ly$\alpha$ line morphology such as the escape fraction, the FWHM,
%the maxima position, also we see that shape of the outgoing spectra
%depends on the position of the observer. It is important to see if
%rotational effects are detectable in observational methods involving
%the Ly$\alpha$ line.

%One of the most used methods to detect high redshift galaxies using
%the Ly$\alpha$ line is using a narrowband selection, we make this
%analysis based on the results obtained by Steidel (2011) in this work
%(EXPLAIN a lit of bit more about their work) they used tree narrowband
%filters for tree different redshifts resumed in
%Table~\ref{table:NBfilters}, We want to know how much the integrated
%flux change due to rotational effects in this NB filters, for the
%models we simulated with CLARA.

%\begin{table*}
%\begin{center}
%\begin{tabular}{cccc}\hline
%Model & SSA22a 4980/80   & HS1549 4667/80 & HS1700 4018/90\\
%\hline
%\end{tabular}
%\caption{
%Fluxes for tree different narrow band filters.
%} 
%\label{table:NBfilters}
%\end{center}
%\end{table*}

%In table \ref{table:NBfilters} we present the results of the flux in
%every  narrowband filter for the homogeneous model at different
%velocities and hydrogen  optical depth $\tau_{0}$. As is well known an
%increase in the optical depth makes  the line peaks separation
%bigger. In fact for some cases the line width is larger  than the NB
%filer width fig.(put ref fig) for those cases we modify the redshift
%in the available range in order to make the peak maxima match the NB
%filter center.  For all cases we found that as velocity increase the
%flux is less.

%In the case of the dusty model we found the same trend of the flux
%with the velocity  but in this cases the effect is no that strong.  We
%also found some dependency with the viewing angle.

%\begin{figure}
%  \includegraphics[width=0.45\textwidth]{NB7tDifVHOM.png}
% \label{figure:efvsNeufeld}\caption{NB filter located at the center of
%   the maxima, with different redshift}  
%\end{figure}

@@

... Comparison with Verhamme et al. results on the rotation

... Compare with Kulas et al (Figure 3), Rotation on the lyman alpha
line convert double peak profiles into a single one. comments about
rotation with inflows and outflows.

... The results derived in this paper have consequences on the
interpretation of galaxy observations in the Lyman alpha line.

..compare steidel et al (2011)

% taking into account equivalent widths results with narrow band
% filters we can make a criterium in order to select wich galaxie
% could be seen in actual observations.
